Rule-Based Preordering on Multiple Syntactic Levels in Statistical Machine Translation

### 2014-09-24 Paper/paper.tex
.

\address{Institute for Anthropomatics \\
Karlsruhe Institute of Technoloy, Germany \\
{\small \tt utcur@student.kit.edu, yuqi.zhang@kit.edu}
}
%
\begin{document}
\maketitle
%
\begin{abstract}
We propose a novel data-driven rule-based preordering approach, which uses the information of syntax tree and word alignment to reorder the words in source sentences before decoding in a phrase-based SMT system between English and Chinese. The preordering algorithm extracts reordering patterns from multiple levels of the syntax trees of the training data and applies the rules on the source sentences in a similar manner. We've conducted experiments in English-to-Chinese and Chinese-to-English translation directions. Our results show that the approach has led to improved translation quality both when it was applied separately on the baseline or when it was combined with some other reordering approaches. As our reordering approach was used alone, it showed an improvement of $1.61$ in BLEU score in the English-to-Chinese translation direction and an improvement of $2.16$ in BLEU score in the Chinese-to-English translation direction, in comparison with the baseline, which used no preordering at all. As our preordering approach were combined with the short rule \cite{short}, long rule \cite{long} and tree rule \cite{tree} based preordering approaches, it showed an improvement of $0.43$ in BLEU score in the English-to-Chinese translation direction and an improvement of $0.3$ in BLEU score in the Chinese-to-English translation direction. Through the translations that used our preordering approach, we've also found improvements in syntactic structure.
\end{abstract}
%
\section{Introduction}

@@

Most state-of-the-art phrase-based SMT systems use language model, phrase table or decoder to adjust the word order. Or an additional reordering model is used in the log-linear model for word reordering. However, these methods may have some disadvantages, such as some don't handle long-distance reordering, some don't handle unstructured reordering and some are rather time consuming.

Encouraged by the results from the paper \cite{short}, \cite{long} and \cite{tree}, we further propose a new data-driven, rule-based preordering method, which extracts and applies reordering rules based on syntax tree. The method is called Multi-Level-Tree (MLT) reordering, which orders the constituents on multiple levels of the syntax tree all together. This preordering method rearranges the words in source language into a similar order as they are supposed to be in the target language before translation. With the appropriate word order, better translation quality can be achieved. Especially, our preordering method is very suitable for translation between language pairs like English and Chinese, which have very different word orders. Besides, the method can also be combined with the above mentioned rule-based reordering methods to achieve better translation quality.

@@

Rottmann and Vogel 2007 \cite{short} introduced the idea of extracting reordering rules from the POS tag sequences of training data and use them for reordering. Niehues and Kolss 2009 \cite{long} went further, and developed a method for long-distance word reordering, which works good on German-English translation task due to the long-distance shift of verbs. The method extracts discontinuous reordering rules in addition to the continuous ones, which contains a placeholder to match several words and enables the word to shift cross long distance.

@@

The word order between English and Chinese differs very significantly. For one, the words in Chinese have generally different origins as those in English, which leads to very different vocabulary and word construction. Sometimes it is very hard to find corresponding words in the other language. For example, some prepositions in Chinese have very different usage than those prepositions in English. Also the continuous writing of Chinese without spaces makes this problem more severe, since word boundaries are not always so clear in Chinese. The text needs to be segmented first before translation. A word segmentation process is used to separate the words, but the results may not always be ideal.

@@

\begin{itemize}
\item \emph{Pre-modifier instead of post-modifier}\\ In Chinese people tend to use pre-modifier rather than post-modifier. This involves the position change of adverbials, relative clauses and preposition phrases during translation.
\item \emph{Construction of questions}
\item \emph{Special sentence constructions}\\ For example, \emph{B\^{a}}-construction in Chinese and sentence constructions such as \emph{there be} and inverted negative sentences in English don't have correspondence in the other language in general.
\item \emph{Long distance word position change}
\end{itemize}

@@

\begin{itemize}
\item \emph{Short rules} \cite{short} \\
Short rules are extracted based on the sequences of adjacent words or their POS tags in the sentence from training data. Sequences of adjacent words or POS tags are observed, rules are then extracted if the same reordering patterns appear frequently. Following is a reordering pattern of a short rule:\medskip\\
\texttt{WRB MD DT -> DT WRB MD}
\item \emph{Long rules} \cite{long}\\
Long rules are specially designed to improve the long distance word reordering for language pairs like English and German. The rules are based on POS tags of the text, and use a placeholder ($*$) to represent one or more words. Following is a reordering pattern of a long rule:\medskip\\
\texttt{NN $*$ MD -> $*$ MD NN}

\item \emph{Tree rules} \cite{tree}\\
While short rules and long rules are based on the flat structure of sentences, tree rules reorder sentences by using information from sentences' syntactic structure. The syntax tree and word alignment of the training corpus are used to train the reordering rules. The tree rules reorder the words both on the word level and on the constituent level. Following is a reordering pattern of a tree rule:\medskip\\
\texttt{NP ( ADJP JJ NN ) -> JJ NN ADJP}\medskip\\
The parenthesis in the example represents the hierarchy in the syntax tree. The left side of the rule corresponds a tree with the root labeled with \emph{NP} and three children, each labeled with \emph{ADJP}, \emph{JJ} and \emph{NN}. 
\end{itemize}

However, not all these reordering problem of Chinese translation can be solved well by using these reordering rules based either on POS tag sequence or single level of syntax tree. In order to improve the reordering between English and Chinese, we need reordering method that can handle more complicated, unstructural word order change. Inspired by the ideas of reordering on syntax tree and hierarchical phrases, we created the MLT reordering, which reorders words based on multiple syntactic levels and can handle long distance word position change and complicated word position change very well.

@@

\noindent \texttt{NP ( \textbf{CD}$_0$ NP ( NP ( \textbf{JJ}$_1$ \textbf{NNS}$_2$ ) PP ( \textbf{IN}$_3$ \textbf{NP}$_4$ ) ) ) -> NP IN CD JJ NNS}\medskip \\
\texttt{or -> 4 3 0 1 2 (with index)}\bigskip

where the POS tags in bold corresponds to words in the leaf nodes of the syntax tree, which are parts that are actually reordered and have a darker gray color in the figure. The parentheses in the reordering pattern indicate the corresponding hierarchies in the syntax tree. The reordering can be alternatively represented with index, which is the actual internal representation to avoid ambiguity caused by POS tags with the same name and we use this representation throughout this paper.

@@

The results shows increasing scores as we used reordering methods from short rules, long rules, tree rules to MLT rules. And better BLEU scores were achieved as we combined the different reordering rules. The MLT rules achieved better BLEU scores and TER scores in both translation directions, not only when it was used alone, but also it was added to the other reordering rules. As the MLT reordering rules were used alone, it showed an improvement of $1.61$ in BLEU score in the English-to-Chinese translation direction and an improvement of $2.16$ in BLEU score in the Chinese-to-English translation direction, in comparison with the baseline, which used no reordering at all. As the MLT reordering rules were combined with the other existing reordering rules, a further improvement of $0.43$ in BLEU score (from $13.38$ to $13.81$) was shown in the English-to-Chinese translation direction, as well as a further improvement of $0.3$ in BLEU score (from $23.84$ to $24.14$) in the Chinese-to-English translation direction.

>>>>

\address{Institute for Anthropomatics \\
Karlsruhe Institute of Technology, Germany \\
{\small \tt utcur@student.kit.edu, yuqi.zhang@kit.edu}
}
%
\begin{document}
\maketitle
%
\begin{abstract}
We propose a novel data-driven rule-based preordering approach, which uses the information of syntax tree and word alignment to reorder the words in source sentences before decoding in a phrase-based SMT system between English and Chinese. The preordering algorithm extracts reordering patterns from multiple levels of the syntax trees of the training data and applies the rules on the source sentences in a similar way. We've conducted experiments in English-to-Chinese and Chinese-to-English translation directions. Our results show that the approach has led to improved translation quality both when it was applied separately on the baseline or when it was combined with some other reordering approaches. As our reordering approach was used alone, it showed an improvement of $1.61$ in BLEU score in the English-to-Chinese translation direction and an improvement of $2.16$ in BLEU score in the Chinese-to-English translation direction, in comparison with the baseline, which used no word reordering. As our preordering approach were combined with the short rule \cite{short}, long rule \cite{long} and tree rule \cite{tree} based preordering approaches, it showed further improvements of up to $0.43$ in BLEU score in the English-to-Chinese translation direction and further improvements of up to $0.3$ in BLEU score in the Chinese-to-English translation direction. Through the translations that used our preordering approach, we've also found improvements in syntactic structure.
\end{abstract}
%
\section{Introduction}

@@

Most state-of-the-art phrase-based SMT systems use language model, phrase table or decoder to adjust the word order. Or an additional reordering model is used in the log-linear model for word reordering. However, these methods may have some disadvantages, such as some don't handle long-distance reordering, some don't handle unstructured reordering and some are rather time-consuming.

Encouraged by the results from the result from Rottmann and Vogel 2007 \cite{short}, Niehues and Kolss 2009 \cite{long} and Herr\-mann et al. 2013 \cite{tree}, we further propose a new data-driven, rule-based preordering method, which extracts and applies reordering rules based on syntax tree. The method is called Multi-Level-Tree (MLT) reordering, which orders the constituents on multiple levels of the syntax tree all together. This preordering method rearranges the words in source language into a similar order as they are supposed to be in the target language before translation. With the proper word order, better translation quality can be achieved. Especially, our preordering method is very suitable for translation between language pairs like English and Chinese, which have very different word orders. Besides, the method can also be combined with the above mentioned rule-based reordering methods to achieve better translation quality.

@@

Rottmann and Vogel 2007 \cite{short} introduced the idea of extracting reordering rules from the POS tag sequences of training data and use them for reordering. Niehues and Kolss 2009 \cite{long} went further, and developed a method for long-distance word reordering, which works good on German-English translation task due to the long-distance shift of verbs. The method extracts discontinuous reordering rules in addition to the continuous ones, which uses a placeholder to match several words and enables the word to shift cross long distance.

@@

The word order between English and Chinese differs very much. For one, the words in Chinese have generally different origins as those in English, which leads to very different vocabulary and word construction. Sometimes it is very hard to find corresponding words in the other language. For example, some prepositions in Chinese have very different usage than those prepositions in English. Also the continuous writing of Chinese without spaces makes this problem more severe, since word boundaries are not always so clear in Chinese. The text needs to be segmented first before translation. A word segmentation process is used to separate the words, but the results may not always be ideal.

@@

\begin{itemize}
\item \emph{Pre-modifier instead of post-modifier}\\ In Chinese people tend to use pre-modifier rather than post-modifier. This involves the position change of adverbials, relative clauses and preposition phrases during translation.
\item \emph{Construction of questions}\\
The two languages have very different ways to construct questions, which raises word order problems for translating questions.
\item \emph{Special sentence constructions}\\ For example, \emph{B\^{a}}-construction in Chinese and sentence constructions such as \emph{there be} and inverted negative sentences in English don't have correspondence in the other language in general.
\item \emph{Long distance word position change}\\
Word reordering between English and Chinese often involves word shift cross long distance.
\end{itemize}

@@

\begin{itemize}
\item \emph{Short rules} \cite{short} \\
Short rules are extracted based on the sequences of adjacent words or their POS tags in sentences from training data. Sequences of adjacent words or POS tags are observed, rules are then extracted if the same reordering patterns appear frequently. Following is a reordering pattern of a short rule:\medskip\\
\texttt{WRB MD DT -> DT WRB MD}
\item \emph{Long rules} \cite{long}\\
Long rules are specially designed to improve the long distance word reordering for language pairs like English and German. The rules are based on POS tags of the text, and use a placeholder ($*$) to represent one or more words. Following is a reordering pattern of a long rule:\medskip\\
\texttt{NN $*$ MD -> $*$ MD NN}

\item \emph{Tree rules} \cite{tree}\\
While short rules and long rules are based on the flat structure of sentences, tree rules reorder sentences by using information from sentences' syntactic structure. The syntax tree and word alignment of the training corpus are used to train the reordering rules. The tree rules reorder the words both on the word level and on the constituent level. Following is a reordering pattern of a tree rule:\medskip\\
\texttt{NP ( ADJP JJ NN ) -> JJ NN ADJP}%\medskip\\
%The parenthesis in the example represents the hierarchy in the syntax tree. The left side of the rule corresponds a tree with the root labeled with \emph{NP} and three children, each labeled with \emph{ADJP}, \emph{JJ} and \emph{NN}. 
\end{itemize}

However, not all these reordering problems of translating Chinese can be solved well by using these reordering rules based either on POS tag sequence or single level of syntax tree. In order to improve the reordering between English and Chinese, we need reordering method that can handle more complicated, unstructural word order change. Inspired by the ideas of reordering on syntax tree and hierarchical phrases, we created the MLT reordering, which reorders words based on multiple syntactic levels and can handle long distance word position change and complicated word position change very well.

@@

\noindent \texttt{NP ( \textbf{CD}$_0$ NP ( NP ( \textbf{JJ}$_1$ \textbf{NNS}$_2$ ) PP ( \textbf{IN}$_3$ \textbf{NP}$_4$ ) ) ) -> NP IN CD JJ NNS}\medskip \\
\texttt{-> 4 3 0 1 2 (alternative with index)}\bigskip

The POS tags in bold corresponds to words in the leaf nodes of the syntax tree, which are parts that are actually reordered and have a darker gray color in the figure. The parentheses in the reordering pattern indicate the corresponding hierarchies in the syntax tree. The reordering can be alternatively represented with index, which is the actual internal representation to avoid ambiguity caused by POS tags with the same name and we use this representation throughout this paper.

@@

The results shows increasing scores when we used reordering methods from short rules, long rules, tree rules to MLT rules. And better BLEU scores were achieved when we combined the different reordering rules. The MLT rules achieved better BLEU scores and TER scores in both translation directions, not only when it was used alone, but also it was added to the other reordering rules. As the MLT reordering rules were used alone, it showed an improvement of $1.61$ in BLEU score in the English-to-Chinese translation direction and an improvement of $2.16$ in BLEU score in the Chinese-to-English translation direction, in comparison with the baseline, which used no reordering at all. As the MLT reordering rules were combined with the other existing reordering rules, a further improvement of $0.43$ in BLEU score (from $13.38$ to $13.81$) was shown in the English-to-Chinese translation direction, as well as a further improvement of $0.3$ in BLEU score (from $23.84$ to $24.14$) in the Chinese-to-English translation direction.
